Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-08-04 prop_letter.tex
prop.bib @InProceedings{Cabello2017} Added...

The effect of finite precision and imperfections on contextuality and the
Kochen-Specker theorem is shown to have a resolution in terms of our QIVPM framework.  
To proceed, we use a formulation of the ``Mermin-Peres square'' \nb\cite{peres1995quantum,Mermin1990Simple} that uses the
expectation value of a physical observable with respect
to a QIVPM, where the interval-valued
expectation value is consistent with the expectation value of the
same observable in conventional quantum mechanics with infinite resources. We show that, 
as long as errors and uncertainties remain smaller than a particular value, the
Kochen-Specker theorem remains valid. In other words, we established a rigorous bound for 
its validity and, therefore, the limits for physical reality to remain contextual.

A key aspect of a quantum measurement is its statistical nature, as reflected
by application of the Born rule to determine the probability of a given measurement 
outcome \nb\citep{Born1983,peres1995quantum,544199}. The Gleason
theorem establishes that there is no alternative to the Born rule by 
demonstrating, using reasonable continuity arguments in Hilbert spaces
of dimension $d\ge3$, the existence of a unique state 
$\rho$ consistent with the statistical predictions computed from 
that rule \nb\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
It is then legitimate to question what happens when infinite precision
cannot be achieved. We will show that, if the effect of finite precision is incorporated, 
while a QIVPM  might not be consistent
with a unique state\nb$\rho$ on all projectors defined on a Hilbert space of dimension $d\ge3$, 
there is a mathematically precise sense in which one recovers the original Gleason theorem asymptotically. 
Specifically, if for bounded resources the QIVPM is induced by some mixed states \nb$\rho'$, 
those states will converge to a unique state $\rho$ when the size of the intervals characterizing 
the uncertainties tends to zero.

The paper is organized as follows.   In Sec.\ \ref{sec:fuzzy} we begin by introducing the foundations of quantum probability space and fuzzy measurement.  We then move on to introduce the concept of quantum interval valued probability in Sec.\ \ref{sec:Interval-Uncertainty}.  This QIVPM will provide the mathematical framework to quantify the impact of finite precision measurements on quantum mechanics.  For instance, in Sec.\ \ref{sec:Kochen-Specker},  we quantify the domain of validity of a  contextual measurement, thus addressing the conditions under which the Kochen-Specker theorem applies.  This provides a way of resolving the Meyer-Mermin debate, which is the main result of our Theorem \ref{cor:Kochen-Specker-IVPM}.  In Sec.\ \ref{sec:Gleason}, we study Gleason's theorem in the context of imprecise measurements, concluding that QIVPM provides a framework with quantitative bounds in which Gleason's theorem, while formally invalid in a universe with bounded resources, holds asymptotically in a mathematically precise way.  Finally, we present our conclusions in Sec.~\ref{sec:Conclusion}.

>>>>

The effect of finite precision and imperfections on contextuality and the
Kochen-Specker theorem is shown to have a resolution in terms of our QIVPM framework.  
To proceed, we use a formulation of the ``Mermin-Peres square''\nb\cite{peres1995quantum,Mermin1990Simple} that uses the
expectation value of a physical observable with respect
to a QIVPM, where the interval-valued
expectation value is consistent with the expectation value of the
same observable in conventional quantum mechanics with infinite resources. We show that, 
as long as errors and uncertainties remain smaller than a particular value, the
Kochen-Specker theorem remains valid. In other words, we established a rigorous bound for 
its validity and, therefore, the limits for physical reality to remain contextual.

A key aspect of a quantum measurement is its statistical nature, as reflected
by application of the Born rule to determine the probability of a given measurement 
outcome\nb\citep{Born1983,peres1995quantum,544199}. The Gleason
theorem establishes that there is no alternative to the Born rule by 
demonstrating, using reasonable continuity arguments in Hilbert spaces
of dimension $d\ge3$, the existence of a unique state 
$\rho$ consistent with the statistical predictions computed from 
that rule\nb\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
It is then legitimate to question what happens when infinite precision
cannot be achieved. We will show that, if the effect of finite precision is incorporated, 
while a QIVPM  might not be consistent
with a unique state\nb$\rho$ on all projectors defined on a Hilbert space of dimension $d\ge3$, 
there is a mathematically precise sense in which one recovers the original Gleason theorem asymptotically. 
Specifically, if for bounded resources the QIVPM is induced by some mixed states\nb$\rho'$, 
those states will converge to a unique state $\rho$ when the size of the intervals characterizing 
the uncertainties tends to zero.

The paper is organized as follows.   In Sec.\ \ref{sec:fuzzy} we begin by introducing the foundations of quantum probability space and fuzzy measurement.  We then move on to introduce the concept of quantum interval valued probability in Sec.\ \ref{sec:Interval-Uncertainty}.  This QIVPM will provide the mathematical framework to quantify the impact of finite precision measurements on quantum mechanics.  For instance, in Sec.\ \ref{sec:Kochen-Specker},  we quantify the domain of validity of a  contextual measurement, thus addressing the conditions under which the Kochen-Specker theorem applies.  This provides a way of resolving the Meyer-Mermin debate\nb\footnote{Cabello has referred to this as the second contextuality war~\citep{Cabello2017}.}, which is the main result of our Theorem \ref{cor:Kochen-Specker-IVPM}.  In Sec.\ \ref{sec:Gleason}, we study Gleason's theorem in the context of imprecise measurements, concluding that QIVPM provides a framework with quantitative bounds in which Gleason's theorem, while formally invalid in a universe with bounded resources, holds asymptotically in a mathematically precise way.  Finally, we present our conclusions in Sec.~\ref{sec:Conclusion}.
